\speculake: Hiding Malware in Speculative Execution

### 2018-08-04 paper/triggers.tex
figure updates
discussion pass
Alternative triggers

In addition to using separate trigger programs, ExSpectre can also use
\emph{trigger inputs} to a payload program to initiate its malicious behavior.
While existing fuzzers and symbolic execution tools can discover traditional
input triggers, we can leverage other Spectre variants...

TODO: Jack

@@

TODO: Jack

>>>>

\FigSpectreOne

In addition to using separate trigger programs, ExSpectre can also use
\emph{trigger inputs} to a payload program to initiate its malicious behavior.
While existing fuzzers and symbolic execution tools can discover traditional
input triggers, we can leverage other Spectre variants to obfuscate speculative
gadget access.

We make use of Specualtive Buffer Overflows (SBO) to redirect control flow
to a speculative entry point specified by user input. Using Spectre variant 1.1 
we can design a program that takes seemingly arbitrary user input and performs complete 
bounds checks such that evaluated code is safe, however a direct branch can
be bypassed speculatively given significant lead up with consistent input 
to train the branch predictor (see Figure~\ref{fig:spectre-one}). This allows the input
to become the trigger.

A direct branch followed by a store operation with user controlled index and
value can be coerced into overflowing a local array. 
As seen in code snippet from Kiriansky and Waldspurger~\cite{kiriansky2018speculative}:
\begin{lstlisting}
    if (y < lenc)
        c[y] = z;
\end{lstlisting}
with user controlled 
\texttt{y} and \texttt{z} and sporadically uncached \texttt{lenc}, an attacker 
can \textit{speculatively} overflow array \texttt{c}
to overwrite a return address, function pointer, etc. and redirect control flow.

We make use of this in a willing payload such that user input can intentionally 
mistrain the branch predictor. The speculative entry point is also chosen by the 
trigger in this scenario as \texttt{z} contains the address of the 
speculative entry point, allowing the attacker to create a ROP style speculative 
execution path through the payload.

@@

Speculative Store Bypass (SSB) can similarly be used to construct an internal trigger
using race conditions on a store and a subsequent load that the cpu believes 
will not alias but in reality do~\cite{spec-store-bypass}.

To redirect control flow a store to a function pointer or indirect branch 
target pointer could be speculated to a stale value. While the program never 
executes at this stale address in reality, the speculative gadgets have a 
real effect on shared resources. This attack again allows ROP style 
chains to execute speculative gadgets.

The speculative primitive is shared by all versions of Spectre as an upper
bound on what can be done speculatively. What is required to adapt a spectre 
vulnerability into the \speculake technique is an attacker controlled method of 
triggering control flow redirection in a willing payload to a speculative entry point.
